---
title: "p8105_hw2_lz2951"
author: "lz2951"
date: "2023-10-04"
output: github_document
---

## Library required packages

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1

## Clean the data in pols-month.csv

```{r}
pols_month = 
  read_csv(file = "./data/pols-month.csv") |> 
  janitor::clean_names() |>
  separate(mon, into = c("year", "month", "day"), sep = "-", convert = TRUE) |>
  mutate(
    month = case_match(
      month,
      1 ~ "January",
      2 ~ "February",
      3 ~ "March",
      4 ~ "April",
      5 ~ "May",
      6 ~ "June",
      7 ~ "July",
      8 ~ "August",
      9 ~ "September",
      10 ~ "October",
      11 ~ "November",
      12 ~ "December"
    )
  ) |>
  pivot_longer(
    c("prez_dem", "prez_gop"),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "party"
  ) |>
  filter(party == 1) |>
  select(-party, -day)
```

## Clean the data in snp.csv

```{r}
snp = 
  read_csv(file = "./data/snp.csv", col_types = cols(date = col_date(format = "%m/%d/%y"))) |>
  janitor::clean_names() |>
  separate(date, into = c("year", "month", "day"), sep = "-", convert = TRUE) |>
  mutate(year = if_else(year > 2023, year - 100, year)) |>
  arrange(year, month) |>
  mutate(
    month = case_match(
      month,
      1 ~ "January",
      2 ~ "February",
      3 ~ "March",
      4 ~ "April",
      5 ~ "May",
      6 ~ "June",
      7 ~ "July",
      8 ~ "August",
      9 ~ "September",
      10 ~ "October",
      11 ~ "November",
      12 ~ "December"
    )
  ) |>
  select(year, month, close)
```

## Clean the unemployment data

```{r}
unemployment =
  read_csv(file = "./data/unemployment.csv") |>
  janitor::clean_names() |>
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) |>
  mutate(
    month = case_match(
      month,
      "jan" ~ "January",
      "feb" ~ "February",
      "mar" ~ "March",
      "apr" ~ "April",
      "may" ~ "May",
      "jun" ~ "June",
      "jul" ~ "July",
      "aug" ~ "August",
      "sep" ~ "September",
      "oct" ~ "October",
      "nov" ~ "November",
      "dec" ~ "December"
    )
  )
```

## Merge three datasets

```{r}
complete_df =
  left_join(pols_month, snp, join_by("year", "month")) |>
  left_join(unemployment, join_by("year", "month"))
```

# Problem 2

## Import the data in Mr. Trash Wheel sheet and calculate the homes powered

```{r}
MrTrashWheel =
  read_excel(
    path = "./data/202309 Trash Wheel Collection Data.xlsx", 
    sheet = "Mr. Trash Wheel",
    range = "A2:N586"
  ) |>
  janitor::clean_names()
MrTrashWheel$year = as.numeric(MrTrashWheel$year)
```

According to the homes powered note, calculate the amount of homes powered:
Each ton of trash equates to on average 500 kilowatts of electricity.  An average household will use 30 kilowatts per day.

```{r}
MrTrashWheel = 
  mutate(
    MrTrashWheel, 
    homes_powered = weight_tons*500/30
  )
```

## Import the data in Professor Trash Wheel sheet and calculate the homes powered

```{r}
ProfTrashWheel =
  read_excel(
    path = "./data/202309 Trash Wheel Collection Data.xlsx", 
    sheet = "Professor Trash Wheel",
    range = "A2:M108",
  ) |>
  janitor::clean_names() |>
  mutate(
    homes_powered = weight_tons*500/30
  )
```

## Import the data in Gwynnda Trash Wheel sheet and calculate the homes powered

```{r}
GwyTrashWheel =
  read_excel(
    path = "./data/202309 Trash Wheel Collection Data.xlsx", 
    sheet = "Gwynnda Trash Wheel",
    range = "A2:L157",
  ) |>
  janitor::clean_names() |>
  mutate(
    homes_powered = weight_tons*500/30
  )
```

## Combine those three datasets

First, to keep track of which Trash Wheel is which, add an additional variable to all datasets.

```{r}
MrTrashWheel = mutate(MrTrashWheel, trash_wheel = "MrTrashWheel")
ProfTrashWheel = mutate(ProfTrashWheel, trash_wheel = "ProfTrashWheel")
GwyTrashWheel = mutate(GwyTrashWheel, trash_wheel = "GwyTrashWheel")
```

Then remove variables not shared by all three datasets and combine three datasets:

```{r}
shared_variables = 
  intersect(colnames(MrTrashWheel), colnames(ProfTrashWheel)) |>
  intersect(colnames(GwyTrashWheel))
MrTrashWheel = select(MrTrashWheel, all_of(shared_variables))
ProfTrashWheel = select(ProfTrashWheel, all_of(shared_variables))
GwyTrashWheel = select(GwyTrashWheel, all_of(shared_variables))
TrashWheels = bind_rows(MrTrashWheel, ProfTrashWheel, GwyTrashWheel)
```

# Problem 3

## Import, clean, and tidy the dataset of baseline demographics

```{r}
mci_baseline = 
  read_csv(
    file = "./data/data_mci/MCI_baseline.csv", 
    skip = 1
  ) |>
  janitor::clean_names() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "Male",
      0 ~ "Female"
    ),
    apoe4 = case_match(
      apoe4,
      1 ~ "carrier",
      0 ~ "non-carrier"
    )
  )
mci_baseline$age_at_onset[mci_baseline$age_at_onset == "."] = "Inf"
mci_baseline$age_at_onset = as.numeric(mci_baseline$age_at_onset)
mci_baseline = filter(mci_baseline, age_at_onset > current_age)
```

## Import, clean, and tidy the dataset of longitudinally observed biomarker values

```{r}
mci_amyloid = 
  read_csv(
    file = "./data/data_mci/mci_amyloid.csv", 
    skip = 1
  ) |>
  janitor::clean_names() |>
  rename(id = "study_id")
```

## Combine two datasets

```{r}
ad_df = left_join(mci_baseline, mci_amyloid, by = join_by(id))
```

